\chapter{Introduction}\label{sec:Introduction}

To get your \bdq{Coding Guardrails for \Java in a Nutshell}\index{Coding Guardrails for Java in a Nutshell}, you can take them from below:\footnote{obviously stolen from the \bdq{Zen of Python}\autocite{WIKIPEDIA:ZenOfPython,PYTHON_ORG_MAILING_LIST:ThePythonWay}\index{Zen of Python} and slightly amended~…}
\begin{enumerate}[nosep]\label{lst:NutshellGuardrails}
	\item{Beautiful is better than ugly.}\label{lst:ZoP:BeautifulVsUgly}
	\item{Explicit is better than implicit, and verbosity is your friend.}\label{lst:ZoP:ExplicitVsImplicit}
	\item{Do not repeat yourself.}\label{lst:ZoP:DRY}\index{DRY Principle}
	\item{Simple is better than complex.}\label{lst:ZoP:SimpleVsComplex}
	\item{Complex is better than complicated.}\label{lst:ZoP:ComplexVsComplicated}
	\item{Flat is better than nested.}\label{lst:ZoP:FlatVsNested}
	\item{Sparse is better than dense.}\label{lst:ZoP:SparseVsDense}
	\item{Readability counts.}\label{lst:ZoP:Readablity}
	\item{Special cases aren't special enough to break the rules~…}\label{lst:ZoP:SpecialCases}
	\item{…~although practicality beats purity.}\label{lst:ZoP:Practicality}
	\item{Errors should never pass silently and Exceptions may never be swallowed~…}\label{lst:ZoP:ErrorsMayNeverBeSwallowed}
	\item{…~unless \textit{explicitly} silenced.}\label{lst:ZoP:ErrorsAreSilenced}
	\item{\textit{Either} log (and terminate) \textit{or} throw, neither both.}\label{lst:ZoP:LogOrThrow}
	\item{In the face of ambiguity, refuse the temptation to guess.}\label{lst:ZoP:DontGuess}
	\item{There should be one – and preferably only one – obvious way to do it~…}\label{lst:ZoP:OneWay}
	\item{…~although that way may not be obvious at first.}
	\item{Now is better than never~…}\label{lst:ZoP:Now}
	\item{…~although never is often better than \textit{right now}.}\label{lst:ZoP:RightNow}
	\item{If the implementation is hard to explain, it's a bad idea.}\label{lst:ZoP:BadIdea}
	\item{If the implementation is easy to explain, it may be a good idea.}\label{lst:ZoP:GoodIdea}
	\item{Think about your future self and add a comment.}\label{lst:ZoP:FutureSelf}
	\item{Without the unit tests\index{unit test}, the code is not complete~…}\label{lst:ZoP:NotCompleteWithoutTest}
	\item{…~neither it will be complete without documentation.}\label{lst:ZoP:NotCompleteWithoutDocumentation}
	\item{Always do it right in the first place!}\label{lst:ZoP:BasicRule}
	\item{Have Fun!}\label{lst:ZoP:HaveFun}
\end{enumerate}

\input{introduction/Section_TheGoals}
\input{introduction/Section_AboutThisDocument}
\input{introduction/Section_HistoryAndImplementation}
\input{introduction/Section_OtherProgrammingLanguages}
\input{introduction/Section_CodeGenerators}
\input{introduction/Section_CodingForSustaining}
\input{introduction/Section_ToolSupport}
\input{introduction/Section_HowToUseThisDocument}
\input{introduction/Section_HowToImplement}
\input{introduction/Section_AboutTheAuthor}
\input{introduction/Section_TheBasicRule}

%==============================================================================
% End of file!!
%==============================================================================
