\section{The Goals}
The main goal of coding conventions in general is to help you writing better code.

So what is \textit{good code}?

\paragraph{Good Code} does not have bugs and is running fast, without wasting valuable resources.

\bdq{Does not have bugs} also means that the code is doing what it is intended to do. We all know that it is nearly impossible to produce completely bug free code once the program is not damn stupid simple, for several reasons. One important reason is that already the specification, used as the base for the implementation, is often faulty. So the code is doing what it should do \textit{according to the specification}, but it is not doing what it is \textit{really} meant to do~…

However, we should not forget that there is still the possibility for \textit{real} bugs.

And often you see the situation that your code is getting slower and slower, or is eating up more CPU cycles and/or memory than available. That may be due to higher data volumes or more users or lots of other reasons.

But as a result of all these scenarios, you will have to modify your code – although, in most cases, it will be someone else who have to dig into \textit{your} code.  Or you are the unlucky one who needs to revise someone else's code. And if the code is only old enough, it does not really matter if you wrote it or someone else – refer to the \ngref, bullet point \ref{lst:ZoP:FutureSelf}.

So what you want is code that helps you with being modified – you want to have \textit{maintainable code}.

\paragraph{Maintainable Code}\index{maintainable code} supports its modification in various ways.

From a magazine dealing with software development tools, I found the following statement\footnote{I noted down the statement, but forgot to note the source – so if you know that source, I would appreciate if you would share it with me.}:

\begin{quote}
``If you compare software development with the Apollo program, most programmers would be very successful bringing man to the moon, but never will get them back alive because of their common incapability to deliver software that can be maintained over a longer period of time with reasonable costs.'' \\
\textit{Unknown Source}\autocite{UnknownSource1}
\end{quote}

\bdq{Maintainable Code} neither means \bdq{Resusable Code} nor \bdq{Generic Code}. Of course, reusing code (isolate it as a method or function) instead of repeating it (copy and paste to another location) can support maintainability, so can it help to generalise an implementation. But first both alone is not sufficient for maintainable code, and second, especially generalisation can be contraproductive.

Larger applications are usually splitted into parts that are organised as libraries or modules. Following the principle of \bdq{Separation of concerns}\autocite{WIKIPEDIA:SoC}\index{Separation of concerns}\index{SoC|see{Separation of concerns}} when designing these parts is helpful to get maintainable software. But some additional thought should be spent on how generic and resuable these libraries need to be. On the other hand, if the library code is really maintainable, it should not be a big deal to add additional features without breaking existing code.

Maintainability has a lot of facets, and I will discuss several of them in the run of this document. But before you can even think of modifying the code – either for fixing a bug, optimising the memory consumption, or adding new functionality – you need to identify the location.  

For that you have to read the already existing code, and of course, you have to understand what you are reading. So a crucial quality of maintainable code is that it is easy to read and \textit{basically} understandable\footnote{Ok, the requirement should be \bsq{easy to understand}, but that is a relative term. If something is \bsq{easily} understandable depends on the experience level of the reader, and here we talk not only about their coding experience, but also about their familiarity with the problem domain. And you cannot always code for the rookies.}.

No matter whether your code will meet all the other criteria that would make it maintainable: if it is not readable, it cannot really be maintainable. 

\paragraph{Readable Code}\index{readable code} can be easily consumed by a human.

\begin{quote}
\bdq{[…], programs must be written for people to read, and only incidentally for machines to execute.}\\
Herold Abelson et. al.: \textit{Structure and Interpretation of Computer Programs}
\autocite{Sussman:StructureAndInterpretationOfComputerPrograms}
\end{quote}

\begin{quote}
\bdq{Everybody can write code that can be read by a computer, but only good developers will write code that can be read by humans.}\footnote{I agree in general, but also the converse is true: only developers that will write code that can be read by humans are good developers.} \\
Martin G. Fowler: \textit{Refactoring: Improving the Design of Existing Code}\autocite{Fowler:Refactoring}
\end{quote}

\begin{quote}
``\textbf{Reading code is more important than writing code.}
Code is read much more often than it is written. Further, when writing code, we usually have the whole context in our head, and take our time; when reading code, we are often context-switching, and may be in more of a hurry. Whether and how particular language features are used ought to be determined by their impact on future readers of the program, not its original author. Shorter programs can be preferable to longer ones, but shortening a program too much can omit information that's useful for understanding the program. The central issue here is to find the right size for the program such that understandability is maximized.

We are specifically unconcerned here with the amount of keyboarding that’s necessary to input or to edit a program. While concision may be a nice bonus for the author, focusing on it misses the main goal, which is to improve the understandability of the resulting program.'' \\
Stuart W. Marks: \textit{Local Variable Type Inference – Style Guidelines}\autocite{Marks:LocalVariableTypeInference:ReadingIsMoreImportantThanWriting}
\end{quote}

And from my own experience I would confirm that verbosity is your friend!\footnote{Refer to the \ngref, bullet point \ref{lst:ZoP:ExplicitVsImplicit}.}

An external documentation (if existing at all~…) can be helpful to understand the code, in particular regarding the aspects of the problem domain the code is dealing with. But the flow of reading is significantly disturbed when you need to consult that external documentation every second line, same as when you have to look into a dictionary all the time when reading a book in a foreign language. So that external documentation is no substitute for readable code.

Several guidelines and recommendations in this document, in particular regarding how to apply comments and how to write comments, but also those about to write a method, do require significant additional typing. \Java in general has the reputation to be too verbose, and this coding conventions will even add to that verbosity – but as said: that's not generally bad!

But if you are afraid of the keyboarding, you should learn typewriting! For my understanding, someone who does not reach at least 20~WPM\footnote{WPM = \bdq{Words per Minute}} should look for a job outside of software development!\footnote{Just for your comparison: average values for a computer typist are around 41~WPM, the world record is 216~WPM, set by Stella Pajunas in 1946, using an IBM electric typewriter. See also \autocite{RATATYPE}.

According to Google, for the English language WPM roughly translates to CPM (= \bdq{Characters per Minute} or \bdq{Anschläge pro Minute} in German) by multiplying it by 5. The average secretary scores at least 200~CPM, peak values are in the 500 to 900~CPM, with a world record an 955~CPM.}

Writing code is not the only area where you would benefit from mastering that important skill; it will also help you to write all the other stuff you have to deliver in addition to code (documentation, meeting notes, emails, specification documents,~…).

From a site that offers typing courses\autocite{RATATYPE}:

\begin{quote}
\begin{samepage}
``\textbf{WHEN TO START LEARNING}

Developed typing skills can help young people get better study results in school or college, and get better job offers once they have finished school. 65\% of people who learn how to touch type faster are under the age of 24. The main goal of learning to touch type or improving touch typing skills, after age 25 is to become more successful in current job.
\end{samepage}

\begin{samepage}
\textbf{MOTIVATION}

Improving touch typing skills is a must for every person under 18, but everyone can benefit from some time spent practicing. Average typing speed is one of the key skills listed on your resume. Spend some more time practicing and get better results […].''
\end{samepage}
\end{quote}

That they do marketing to sell their product does not make their statements wrong.

Also, modern IDEs do a lot of the typing for you, as long as you take the trouble to configure them appropriately. Refer to chapter \tqfullref{sec:ToolSupport}.

\begin{quote}
``\textbf{Code readability shouldn't depend on IDEs.}
Code is often written and read within an IDE, so it's tempting to rely heavily on code analysis features of IDEs. [Why not relying on that?]

There are two reasons. The first is that code is often read outside an IDE. Code appears in many places where IDE facilities aren’t available, such as snippets within a document, browsing a repository on the internet, or in a patch file. It is counterproductive to have to import code into an IDE simply to understand what the code does.

The second reason is that even when one is reading code within an IDE, explicit actions are often necessary to query the IDE for further information […] This might take only a moment, but it disrupts the flow of reading.

Code should be self-revealing. It should be understandable on its face, without the need for assistance from tools.'' \\
Stuart W. Marks: \textit{Local Variable Type Inference – Style Guidelines}\autocite{Marks:LocalVariableTypeInference:NoIDEDependency}
\end{quote}

\subsection{Conclusion}
To summarize from above, the goal for these coding conventions is to help you to write~…
\begin{itemize}[nosep]
	\item …~good,~…
	\item …~maintainable,~…
	\item …~readable~…
\end{itemize}
…~\Java source code.

As a side effect, all source code looks familiar to all members of the team. That conformity further supports readability, but such a common look also identifies the source code as written by the team, as its trademark.

%==============================================================================
% End of file!!
%==============================================================================
