\section{History and Implementation}
Around the year~2000 I was asked to compile a set of coding guardrails for a team of developers that had been new to Java (and most were new to programming, too). That was the very first version of this document, based on the “Code Conventions for the Java\textsuperscript{TM} Programming Language”\autocite{SUN_CODE_CONVENTIONS} and the way I wrote Java code at that time.

Later versions were created for other developer teams, and several software projects had been delivered successfully working with these coding conventions.

Some of the code from these projects is still alive, proving the initial statement that coding conventions do help to create maintainable software.

I wrote a lot of code using the rules, guidelines and recommendations from this document (and its predecessors, of course), including some libraries that I published on GitHub. Some of the code from these libraries is mentioned in chapter \tqfullref{sec:CodingGuidelines}, starting on page~\pageref{sec:CodingGuidelines}, I also added some stuff as examples in chapter \tqfullref{sec:Examples} in the “\nameref{sec:Appendices}” chapter. If you are interested in the whole package, you should have a look to \autocite{TQUADRAT_ORG}.

\section{Other Programming Languages}\label{sec:OtherProgrammingLanguages}
Of course coding conventions like those described with this document are not only useful for Java programs, but for code in any other programming language also.\footnote{Even for descriptive languages like XML and HTML they make sense, also for documents written in \TeX/\LaTeX.} But each language has its unique features and specialities, meaning that coding conventions written for one programming language do not necessarily match the requirements for another.

Obviously there are some common rules that are valid for every programming language (e.g. the request to use meaningful names), but usually the differences will outweigh the similarities. So it is not a good idea to do something in language \verb#A# just and only it is done that way in programming language \verb#B#. On the other side, it should always be proved whether something that worked fine for one programming language would not be a great idea to be applied to code written in another language, too.

In this sense, I stole the “Zen of Python”\autocite{WIKIPEDIA:ZenOfPython,PYTHON_ORG_MAILING_LIST:ThePythonWay}\index{Zen of Python}\footnote{I added the ellipsis.} and used it for this document:

\begin{itemize}[nosep]
	\item Beautiful is better than ugly.
	\item Explicit is better than implicit.
	\item Simple is better than complex.
	\item Complex is better than complicated.
	\item Flat is better than nested.
	\item Sparse is better than dense.
	\item Readability counts.
	\item Special cases aren't special enough to break the rules~…
	\item …~although practicality beats purity.
	\item Errors should never pass silently~…
	\item …~unless explicitly silenced.
	\item In the face of ambiguity, refuse the temptation to guess.
	\item There should be one – and preferably only one – obvious way to do it~…
	\item …~although that way may not be obvious at first unless you're Dutch.
	\item Now is better than never~…
	\item …~although never is often better than \textit{right now}.
	\item If the implementation is hard to explain, it's a bad idea.
	\item If the implementation is easy to explain, it may be a good idea.
	\item Namespaces are one honking great idea – let's do more of those!
\end{itemize}

Only the last one, regarding “Namespaces”, does not work for code written in Java, unless you translate “Namespace” to “Package” (or “Module”~…).

A good example for the other way round – other programming languages adopting a feature from Java – is the idea to add documentation comments to the source code and to use a tool like JavaDoc\index{javadoc} to externalise them and publish them as the API documentation. It was adopted for various other languages, including so different specimen as C\index{C}/C++\index{C++}, JavaScript\index{JavaScript} and PL/SQL\index{PL/SQL}.\footnote{Although I have to confess that the commenting style does not get part of the specification of these languages, instead it is only supported by external tools; one of these tools is doxygen\autocite{DOXYGEN_HOMEPAGE}\index{doxygen}.} Obviously, if you want to benefit from that adopted feature, you have to write your C++\index{C++} (JavaScript\index{JavaScript}, PL/SQL\index{PL/SQL},~…) code (more precise, your documentation comments) according to coding conventions similar to those defined here.

A sample for code conventions for JavaScript\index{JavaScript} can be found in the web at \autocite{Crockford:JAVASCRIPT_CODE_CONVENTIONS}.

Coding conventions for C\#\index{C\#} are part of Microsoft's documentation for C\#\autocite{MICROSOFT:CsharpDocumentation}, another source can be found on GitHub\autocite{Taranov:CsharpCodingStandards}.

The number of documents dealing with coding standards, naming conventions and style guides for C++\index{C++} is close to infinite. A good starting point could be the ISO page\autocite{ISO:CppCodingStandards}, another one would be Google's C++ Style Guide\autocite{Google:CppStyleGuide}.

\section{Code Generators}
Java code that is automatically generated by another piece of software should implement this coding conventions in the same way as “hand crafted” code. Most generators are highly configurable and/or allow to use code templates that are customisable.\footnote{Perhaps you want to have a look to \autocite{TQUADRAT_ORG_FOUNDATION_JAVACOMPOSER}.}

Especially if base classes will be generated it is important that proper comments are generated at least for all API elements; this covers all \lstinline|public| and \lstinline|protected| elements, and in some cases all package-local elements, too.

Java~6 introduced an annotation \lstinline|javax.annotation.Generated|\index{javax.annotation!Generated} (\lstinline|@Generated|) that should be used to mark generated code. Refer to \autocite{ORACLE_DOC_GENERATED_ANNOTATION} for the details.\footnote{In case you still have to use Java~5 as the source code version, you should consider to create your own \lstinline|@Generated| annotation along the lines given by that existing implementation in Java~6 and later.}

\section{Coding for Sustaining}\label{sec:CodingForSustaining}
And what to do with “legacy” code that has to be fixed, amended or extended? 

Most probably that code will not follow \textit{these} coding conventions (if any at all~…), and it is rarely a good idea to reformat or completely rewrite the existing source code just to align it with the current recommendations and guidelines. Instead a fix should be limited to that area in the source code that is broken.

But if possible, your fix should incorporate the standards defined by this document. Be creative how to implement these standards. So if a method has to be reimplemented completely, format it as described here and add the appropriate JavaDoc comments\footnote{Even when there are no other JavaDoc comments in the whole file~…}, if missing. Or if a new field has to be added to a class, add the “\verb#m_#” prefix to its name (see chapter \tqfullref{sec:Fields}).

Obviously, completely new interfaces and classes should be implemented fully compliant to this conventions.

In this context, see also chapter \tqfullvref{sec:MaintenanceComments}.

\section{Tool Support}\label{sec:ToolSupport}
Several of the rules, guardrails and recommendations in this document can be implemented through the configuration of the programming enviroment that is used, especially those from chapter \tqfullref{sec:FormattingTheSourceCode} that deals with the formatting of the source code.

The chapter \tqfullref{sec:IDEConfiguration} provides some configuration samples for Eclipse and JetBrains IntelliJ IDEA.

\section{How to use this Document?}\label{sec:HowToUseThisDocument}
First, you can use this document as is! It should work for you, as its predecessors already worked for a bunch of project teams.

But you can also shape it closer to your particular needs, if you want! As said, you find the sources for this document on GitHub. Clone the repository on GitHub \autocite{TQUADRAT_ORG_DOCUMENT_REPOSITORY}, replace the examples and the references to my libraries by some that fits better to your project and/or your company, then adjust the other stuff accordingly.

In particular, for your project, you should elaborate on the following topics:
\begin{itemize}
	\item{Chapters \tqfullvref{sec:Projects}, \tqfullvref{sec:Modules}, \tqfullvref{sec:Packages} and \tqfullvref{sec:LibraryFiles}\\
	In no particular order:
	    \begin{itemize}
		    \item{Specify the name of your project.}
		    \item{Specify a format for the names of the library file(s) produced by your project.}
		    \item{Specify the names for the modules and packages of your project.}
	    \end{itemize}}
	
	\item{Chapters \tqfullvref{sec:Logging} and \tqfullvref{sec:LoggerConfiguration}:
	    \begin{itemize}
		    \item{Specify the type of the logging framework you use in your product – if any – or if you want to code against SLF4J (or if you do not want to log at all).}
		    \item{Specify the names of the loggers.}
		    \item{Specify the format for the log messages, in particular for the log levels \verb#DEBUG# and \verb#TRACE# (or \verb#FINE#, \verb#FINER# and \verb#FINEST#, if you are using JDK~Logging).}
		    \item{Provide a (default) logging configuration.}
		    \item{Optionally, you may remove all references to the logging frameworks and APIs that you do \textit{not} use in your project.}
	    \end{itemize}}
	    
	\item{Chapter \tqfullvref{sec:TheNamingDictionary}:
	    \begin{itemize}
		    \item{Add verbs that are specific to your project, remove obsolete/unwanted ones.}
		    \item{Adjust the description for verbs where appropriate.}
		    \item{Do the same for the class name suffixes.}
		    \item{Add a Glossary that explains special terms used in the context of your project (the “Domain Specific Terminology”).}
	    \end{itemize}}
	
	\item{Chapter \tqfullvref{sec:IDEConfiguration}:
	    \begin{itemize}
		    \item{Remove the configuration setting samples for the IDEs that you are not using.}
		    \item{Adjust the configuration setting sample according to the changes you made elsewhere.}
		    \item{Add matching configuration setting samples for additional IDEs used in your environment.}
	    \end{itemize}}
	
	\item{Provide code snippets for recurring patterns that are specific to your project.}
	
	\item{Remove this chapter \tqfullref{sec:HowToUseThisDocument} and the chapter \tqfullref{sec:HowToImplement}.}

\end{itemize}

\section{How to implement these ‘Coding Conventions’ in your Project?}\label{sec:HowToImplement}
First any coding conventions are better than none! So if your team is already adhering to some coding conventions, the first hurdle had been taken already.

Adjust this document that it will incorporate the already accepted guidelines and recommendations; this may require that you remove conflicting recommendations and guidelines – don't care, just do it!

The biggest blocker in this case could be that too much code needs to be modified to match with the coding conventions, meaning tedious work no one likes to do. In this regard, you should also have a look to the chapter \tqfullvref{sec:CodingForSustaining}.

But the most simple case is a new team starting over with a new project from scratch. Just present the document as is~…

Experienced programmers know about the benefits of having detailed guard\-rails they can follow – in particular when they know that the other team members will also follow the same guardrails.

Nevertheless, you may have one or the other hacker in your team that complains that this kind of guardrails are suffocating their creativity.

Perhaps this helps: I was always interested in fotography, and therefore I attended some classes about the topic. There the question was raised what are bad pictures, good pictures and exceptionally artistic pictures, and the professor explained it like this: Fotography has a whole bunch of rules on how to compose a good picture, and you do not even need to know one of these rules to distinguish a bad shot from a good fotography: the bad ones will break the rules.

But the really artistic pictures always break \textit{one} or \textit{the other} rule, but \textit{with intention}. The fotographer's creativity is to choose the rule to break and how – but this first means that they have to know each of the rules and how to follow them (usually referred to as ‘knowing the craft’), and second, that there is a visible intention.

But in most cases the client just wants to get only good pictures.

Same in software development: only few clients pay for a piece of art; instead they expect good craftsmanship.

If this explanation does not help, and you are in the proper position: replace the hacker.

\section{About the Author}
My name is Thomas Thrien, I was born and raised in Germany, still live there, so my primary language is German, not English. You may notice that one or the other time when you read this text: although I have to speak English most of time on the job, I too often fall back to Denglish~…

I am in the IT business since 1982 now, mainly in roles dealing with support, field service, and sustaining, but over the years I also did some development work, and even some IT consulting. I did the ground work with my fingers on the keyboard as well as leading teams of various size, even running whole departments.

But in support and sustaining, you are mainly on the “receiving end” of the development process, suffering from the output of development teams that had never heard about the basics, or don't care about them. This experience was my main motivation to compile this document, and to keep it up to date.
 
If you have questions or comments regarding this document, please send me an email to \href{mailto:thomas.thrien@tquadrat.org}{thomas.thrien@tquadrat.org}. Or leave it at \url{https://tquadrat.github.io/}.

\section{The Basic Rule}\label{sec:TheBasicRule}
One rule should be followed before any other rule or recommendation given in this document, that one listed as bullet point~\ref{lst:ZoP:BasicRule} in the nutshell version of the \hyperref[lst:NutshellGuardrails]{“Coding Guardrails for Java”}:\footnote{And it does not contradict with bullet points~\ref{lst:ZoP:Now} and \ref{lst:ZoP:RightNow}!}

\begin{center}
	\begin{huge}
		\fbox{\textbf{Always do it right in the first place!}}
	\end{huge}
\end{center}

Experiences with lots of programming projects have shown that programmers seldom touch their code again, once it is written.\footnote{Usually, they will return to the code only when it is broken – and then they want to fix the bug as fast as possible, not adding missing comments or reformatting the code~…} As a result, missing comments will never be added, shady programming patterns would not be fixed – with the consequence that in case of a bug the maintainer is lost in a poorly documented chaos of badly formatted source code.

\textit{And this is exactly what we want to avoid!} 

This means there is never ever any excuse for omitting comments, using ‘temporary’ names or doing other ‘funny’ things. And we all know about all the quick and dirty PoCs that made it into a product without significant changes to the code~… – treat a PoC in the same way as product code.\footnote{I was inclined to add “There is nothing like a \textit{disposable prototype}.” to the \ngref~…}

Of course you can interpret this also as “Don't make mistakes”, but we all now that humans make mistakes, and that it is unavoidable that our code will contain one or the other bug. So it means that you should \textit{do your best} to avoid mistakes, and the rules, guidelines and recommendation from this document will help you with that.

So again:

\begin{center}
	\begin{huge}
		\fbox{\textbf{Always do it right in the first place!}}
	\end{huge}
\end{center}

But the most important rule from the \ngref{} is that with the bullet point number~\ref{lst:ZoP:HaveFun}:

\begin{center}
	\begin{huge}
		\fbox{\textbf{Have fun!}}
	\end{huge}
\end{center}

%==============================================================================
% End of file!!
%==============================================================================
