\section{The Basic Rule}\label{sec:TheBasicRule}
In engineering, it is common lore that nothing lives as long as a temporary solution. This is true for software development, too. And the IT business has also a specific variant of this: \bdq{Nothing lives as long as a PoC\index{Proof of Concept}\index{PoC|see{Proof of Concept}}!}

This leads us to the one rule that should be followed before any other rule or recommendation given in this document, that one listed as bullet point~\ref{lst:ZoP:BasicRule} in the nutshell version of the \ngref:\footnote{And it does not contradict with bullet points~\ref{lst:ZoP:Now} and \ref{lst:ZoP:RightNow}!}

\begin{center}
	\begin{huge}
		\fbox{\textbf{Always do it right in the first place!}}
	\end{huge}
\end{center}

Experiences with lots of programming projects have shown that programmers seldom touch their code again, once it is written.\footnote{Usually, they will return to the code only when it is broken – and then they want to fix the bug as fast as possible, not adding missing comments or reformatting the code~…} As a result, missing comments will never be added, shady programming patterns would not be fixed – with the consequence that in case of a bug the maintainer is lost in a poorly documented chaos of badly formatted source code.

\textit{And this is exactly what we want to avoid!} 

This means there is never ever any excuse for omitting comments, using \bsq{temporary} names or doing other \bsq{funny} things. It is not an excuse that it is only a tempory solution, an emergency fix or a last minute Proof of Concept\index{Proof of Concept}.

We all know about all the quick and dirty PoCs\index{Proof of Concept} that made it into a product without significant changes to the code~… – it means that you must treat a PoC in the same way as you would write product code.\footnote{I was inclined to add \bdq{There is nothing like a \textit{disposable prototype}.} to the \ngref~…}

Of course you can interpret this also as \bdq{Don't make mistakes}, and you are not far away from the truth. But we all now that humans make mistakes, and that it is unavoidable that our code will contain one or the other bug. So it means that you should \textit{do your very best} to \textit{avoid} mistakes, and the rules, guidelines and recommendation from this document will help you with that – if you follow them!

So again:

\begin{center}
	\begin{huge}
		\fbox{\textbf{Always do it right in the first place!}}
	\end{huge}
\end{center}

But the most important rule from the \ngref is that with the bullet point number~\ref{lst:ZoP:HaveFun}:

\begin{center}
	\begin{huge}
		\fbox{\textbf{Have fun!}}
	\end{huge}
\end{center}

%==============================================================================
% End of file!!
%==============================================================================
