\section{How to use this Document}\label{sec:HowToUseThisDocument}
First, you can use this document as is! It should work for you, as its predecessors already worked for a bunch of project teams before.

But you can also shape it closer to your particular needs, if you want! Feel free to strip it from parts you do not like, add stuff you think its missing, or change areas you don't like! If you fix errors, I would appreciate some feedback~…

As said, you find the sources for this document on GitHub\index{GitHub}. Clone the repository on GitHub \autocite{TQUADRAT_ORG_DOCUMENT_REPOSITORY}, replace the examples and the references to my libraries by something that fits better to your project and/or your company, then adjust the other stuff accordingly.

In particular, for your project, you should elaborate on the following topics:
\begin{itemize}
    \item{Chapter \tqfullvref{sec:FormattingEmbeddingCode}:
        \begin{itemize}
        	    \item{If necessary, adjust the provided formatting rules for embedded code according to your needs.}
        	    \item{Add missing formatting rules for additional code types.}
        \end{itemize}}
	\item{Chapters \tqfullvref{sec:Projects}, \tqfullvref{sec:Modules}, \tqfullvref{sec:Packages} and \tqfullvref{sec:LibraryFiles}\\
	In no particular order:
	    \begin{itemize}
		    \item{Specify the name of your project.}
		    \item{Specify a format for the names of the library file(s) produced by your project.}
		    \item{Specify the names for the modules and packages of your project.}
	    \end{itemize}}
	
	\item{Chapters \tqfullvref{sec:Logging} and \tqfullvref{sec:LoggerConfiguration}:
	    \begin{itemize}
		    \item{Specify the type of the logging framework you use in your product – if any – or if you want to code against SLF4J (or if you do not want to log at all).}
		    \item{Specify the names of the loggers.}
		    \item{Specify the format for the log messages, in particular for the log levels \verb#DEBUG# and \verb#TRACE# (or \verb#FINE#, \verb#FINER# and \verb#FINEST#, if you are using JDK~Logging).}
		    \item{Provide a (default) logging configuration.}
		    \item{Optionally, you may remove all references to the logging frameworks and APIs that you do \textit{not} use in your project.}
	    \end{itemize}}
	    
	\item{Chapter \tqfullvref{sec:TheNamingDictionary}:
	    \begin{itemize}
		    \item{Add verbs that are specific to your project, remove obsolete/unwanted ones.}
		    \item{Adjust the description for verbs where appropriate.}
		    \item{Do the same for the class name suffixes.}
		    \item{Add a Glossary that explains special terms used in the context of your project (the \bdq{Domain Specific Terminology}).}
	    \end{itemize}}
	
	\item{Chapter \tqfullvref{sec:IDEConfiguration}:
	    \begin{itemize}
		    \item{Remove the configuration setting samples for the IDEs that you are not using.}
		    \item{Adjust the configuration setting sample according to the changes you made elsewhere.}
		    \item{Add matching configuration setting samples for additional IDEs used in your environment.}
	    \end{itemize}}
	
	\item{Provide code snippets for recurring patterns that are specific to your project.}
	
	\item{Remove this chapter \tqfullref{sec:HowToUseThisDocument} and the chapter \tqfullref{sec:HowToImplement}.}

\end{itemize}

The repository contains also a build script \verb#build.sh# that generates the final PDF.

%==============================================================================
% End of file!!
%==============================================================================
