\part{Formatting the Source Code}\label{sec:FormattingTheSourceCode}
As I said already in the introduction, coding conventions are primarily guidelines that should help making code better to maintain. Usually, this starts with making the code better to read.

In this regard, a source file is like any other text document: it benefits from having a clear structure. And that structure is supported by the formatting of the text.

An article may have sections with headlines in bold face, the paragraphs are separated by blank lines and the first lines of them are indented by 0.3~cm, and there are rules regarding line length, hyphenation, and how to treat widows and orphans.

This part of the document deals with the proper formatting of \Java source code files and their structure with the goal to create human readable \textit{documents}.

\begin{quote}
``[…], we want to establish the idea that a computer language is not just a way of getting a computer to perform operations but rather that it is a novel formal medium for expressing ideas
about methodology.'' \\
Herold Abelson et. al.: \textit{Structure and Interpretation of Computer Programs}
\autocite{Sussman:StructureAndInterpretationOfComputerPrograms}
\end{quote}

\subsubsection{Principles}
\begin{enumerate}[label=P\arabic*.]
	\item{The structure of the program text (the source code) must support the flow of reading.
	
	As said in the \ngref, bullet point~\ref{lst:ZoP:Readablity}: \bdq{Readability counts}.}
	
	\item{The formatting of the program text has to be consistent.
	
	Permanent changes in the way things are presented will confuse the reader.}
	
	\item{Navigation through the source file should be easy.
	
	When loaded inside an IDE, navigating inside a Java source file is easy: click on the name of the respective program element and the cursor is moved to the corresponding line in the file.
	
	But when using a plain text editor, the search function is often the only tool you have. A consistent structure with headlines and a defined order would help here, too.}
	
	\item{It does not harm when the source code also looks beautiful.
	
	Bullet point~\ref{lst:ZoP:BeautifulVsUgly} from the \ngref~ is also about the visual appearance of the source code. We can discuss how useful it is to use ASCII art in a source file, but not that it can contribute to the beauty of the program text.}
\end{enumerate}

\input{formattingCode/Chapter_LinesAndSpaces}
\input{formattingCode/Chapter_JavaSourceFileStructure}
\input{formattingCode/Chapter_Declarations}
\input{formattingCode/Chapter_Statements}
\input{formattingCode/Chapter_EmbeddedCode}

%==============================================================================
% End of file!!
%==============================================================================
