\section{Embedded Code}\label{sec:EmbeddedCode}
Sometimes, it is necessary to embed code from another (programming) language into the \Java source code\footnote{If you write a code generator tool, this could even by Java code.}.

First, only relativly small fragments of foreign code should be embedded into the Java source; larger fragments and full documents can be handled better when provided as resources (refer to \tqfullvref{sec:ResourcesClasspath}). I consider something around 15 to 20~lines still as \bsq{small}, but this can be discussed for sure.

Next, the code should be placed into a text block\index{text block} (see \autocite{ORACLE_DOC_LANGUAGE_SPECIFICATION:TextBlocks} for the definition):
\begin{lstlisting}[numbers=left]
final var code = """
````<code1>
````    <code2>
````""";
\end{lstlisting}

The left margin of the text inside the text block\index{text block} is indicated by the position of the closing three double quotes. In the sample above, \verb#<code1># in line~2 is not indented at all, while \verb#<code2># in line~3 is indented by four spaces. The indentation\index{indentation} within the source code does not matter here; it is just important for optical reasons. The \lstinline|" "| indicates the indentation blanks for the string only.

Aligning the opening and the closing delimiters (the three double quotes) is not recommended. This requires reindentation of the text block\index{text block} if the variable name or modifiers are changed.

According to chapter \tqfullref{sec:IndentationCharacter} only blanks are used for the indentation\index{indentation!character} of the Java source code. But it is of course possible that the embedded code requires tabs for its indendation. In this case, you have to the use the respective escape sequence \bdq{\lstinline|\\t|}:
\begin{lstlisting}[numbers=left]
final var codeJavaScript = """
````\tsetTimeout(function() {
````\t\tmyDisplayer("Hello!");
````\t}, 3000);
````""";
\end{lstlisting}

\bdq{Programmer's Guide to Text Blocks} \autocite{Laskey:TextBlocks} is a comprehensive guide to text blocks\index{text block} in \Java. Text blocks will also be discussed more generally again in chapter \tqfullvref{sec:TextBlocks}.

\subsubsection{Principles}
\begin{enumerate}[label=P\arabic*.]
\item{The opening three double-quotes (the \bsq{\textit{delimiter}}) for a text block\index{text block} should be placed at the right end of the previous line, and the closing delimiter should be placed on its own line, at the left margin of the text block.

Use the \bdq{\lstinline|\|} (line-terminator) escape sequence if the trailing line-feed is not wanted.}
\item{Avoid aligning the opening and closing delimiters and the text block's left margin.}
\item{If tabs are needed for the indentation of the code inside the text block, use the escape sequence \bdq{\lstinline|\\t|}. 

Mixing white space will lead to a result with irregular indentation.}
\item{Do not embed large code blocks.

Use resources for extensive code.}
\end{enumerate}

%------------------------------------------------------------------------------

\subsection{Formatting SQL inside Java}\label{sec:FormattingSQLInsideJava}

%------------------------------------------------------------------------------

\subsection{Formatting XML inside Java}\label{sec:FormattingXMLInsideJava}
You should embed only small fragments of an XML document into the Java source; larger fragments and full documents can be handled better when provided as resources.

%------------------------------------------------------------------------------

\subsection{Formatting HTML inside Java}\label{sec:FormattingHTMLInsideJava}
Same as for XML, also only small HTML fragments should be embedded into the Java source code. Anything else should go into a resource file.


%==============================================================================
% End of file!!
%==============================================================================
