\subsection{\bsq{for} Statements}\label{sec:ForStatements}
Since Java~5 the language knows two distinct types of \lstinline|for|\index{for} loops, both described in \autocite{ORACLE_DOC_LANGUAGE_SPECIFICATION:For}.

The basic \lstinline|for| statement\index{for!basic} looks like this:
\begin{lstlisting}
for( <initialization>; <condition>; <update> )
{
    <statements>;
}

for( <initialization>; <condition>; <update> ) <singleStatement>;
\end{lstlisting}
For the details of this basic form, refer to \autocite{ORACLE_DOC_LANGUAGE_SPECIFICATION:ForBasic}.

An enhanced \lstinline|for| statement\index{for!enhanced} (defined in \autocite{ORACLE_DOC_LANGUAGE_SPECIFICATION:ForEnhanced}) would be written as below:

\begin{lstlisting}
for( <declaration> : <iterable> )
{
    <statements>;
}

for( <declaration> : <iterable> ) <singleStatement>;
\end{lstlisting}

The use of blanks, round brackets, curly braces and linefeeds was already discussed in \tqfullvref{sec:WhiteSpace}.

Curly braces can be omitted only when there is just a single statement to be executed in the loop. In addition, this single statement has to be written completely into the same line as the \lstinline|for| itself. In any other case, the curly braces are mandatory.

\bdq{Single statement} translates to \bdq{simple statement} in this context; the samples below are invalid because the loop body statements are not \textit{simple}:
\begin{lstlisting}[numbers=left]
// DISCOURAGED
for( var i = 0; i < max; ++i ) if( v [i] > 0 ) sum += v [i];
for( final var s : texts ) if( !s.empty() ) out.println( s );

// AVOID!!!
for( var i = 0; i < max; ++i ) if( v [i] > 0 )
{
    sum += v [i];
}
else
{
    sum -= v [i];
}

for( final var s : texts ) if( !s.empty() )
{
    System.out.print( "Value: " );
    System.out.println( s );
}
\end{lstlisting}
In particular the samples in the lines~6 and below are very problematic as they are difficult to read and to understand.

The forms
\begin{lstlisting}
// DISCOURAGED
for( <initialization>; <condition>; <update> )
    <statement>;
\end{lstlisting}
and
\begin{lstlisting}
// DISCOURAGED
for( <declaration> : <iterable> )
    <statement>;
\end{lstlisting}
are syntactically correct, but should be avoided as they are error prone in case the code will be changed later.

When explicitly using an iterator\index{iterator}\index{java.util!Iterator}\autocite{ORACLE_DOC_ITERATOR_INTERFACE} with the (basic) \lstinline|for|\index{for!basic} loop, this should look like this:
\begin{lstlisting}
for( final var i = source.iterator(); i.hasNext(); )
{
    final var currentValue = i.next();
    …
}
\end{lstlisting}

An empty loop body\index{for!empty loop} can be written like this:
\begin{lstlisting}[numbers=left]
for( <initialization>; <condition>; <update> )
{
    // Deliberately empty
}

for( <initialization>; <condition>; <update> );
\end{lstlisting}

In case the longer form (lines~1 to 4) is used, the comment is mandatory, because empty code blocks are forbidden. I recommend the shorter forms (line~6).

In an empty \lstinline|for| loop\index{for!empty loop} all the work is done in the initialization, condition, and update clauses.\footnote{In most cases, you should consider to use a \lstinline|while|\index{while} loop instead of an empty \lstinline|for|\index{for} loop. See \hyperref[sec:WhileStatements]{there}.} Therefore an enhanced \lstinline|for| loop\index{for!enhanced} loop without body does not make much sense. It may be possible to write an implementation of \lstinline|java.lang.Iterable|\index{java.lang!Iterable}\autocite{ORACLE_DOC_ITERABLE_INTERFACE} with an iterator\index{java.util!Iterator}\autocite{ORACLE_DOC_ITERATOR_INTERFACE} that does something as a side effect of \lstinline|hasNext()|\index{java.util!Iterator!hasNext()} or \lstinline|next()|\index{java.util!Iterator!next()}, but this would break the contract of the interface and is therefore strongly discouraged.

An otherwise empty loop should not be used to consume CPU cycles; there are methods fitting much better to this purpose.

When using the comma operator in the initialization or update clause of a \lstinline|for|\index{for} statement, avoid the complexity of using more than three variables. If needed, use separate statements before the \lstinline|for|\index{for} loop (for the initialization clause) or at the end of the loop (for the update clause).

Examples:
\begin{lstlisting}[numbers=left]
for( var i = 0, j = target.length; i < source.length && j >= 0; ++i, --j )
{
    target [j] = source [i];
}

var proceed = true;
var i = source.iterator();
for( var j = 0; proceed; )
{
    target [j++] = i.next();
    proceed = i.hasNext() && j < target.length;
}
\end{lstlisting}

When the loop body is larger, you should consider to use a label\index{label}, also when you use \lstinline|break|\index{break} or \lstinline|continue|\index{continue} inside the loop body. The label will then be repeated as the end comment of the code block. This is described in chapter \tqfullref{sec:LabelsAndBreakStatements} and also in  \tqfullref{sec:TrailingOrEndOfLineComments} and \tqfullref{sec:CommentsWhen}.
\begin{lstlisting}[numbers=left]
<Label>: for( <initialization>; <condition>; <update> )
{
    if( <condition1> ) continue <Label>;
    if( <condition2> ) break <Label>;
    <statements>
}   //  <Label>:

<Label>: for( <declaration> : <iterable> )
{
    if( <condition1> ) continue <Label>;
    if( <condition2> ) break <Label>;
    <statements>
}   //  <Label>:
\end{lstlisting}

\subsubsection{Principles}
\begin{enumerate}[label=P\arabic*.]
\item{No blanks between the \lstinline|for|\index{for} keyword and the opening round bracket. The blanks after the semi-colons are mandatory.

Refer to \tqfullvref{sec:ParameterParenthesis} on how to place the white space.}
\item{If the loop body is a simple single statement, it can be written without curly braces to the same line as the \lstinline|for|\index{for} itself.

Otherwise curly braces are mandatory.}
\item{The curly braces are on lines of their own, with the same indentation level as the \lstinline|for|\index{for} keyword. The code of the body will be indented one level.}
\item{Do not use a \lstinline|for|\index{for} loop to consume CPU cycles only!}
\item{Empty \lstinline|for|\index{for!empty loop} loops do all the work in the initialisation and/or update clauses.

Prefer a \lstinline|while|\index{while} loop over an empty \lstinline|for|\index{for!empty} loop.}
\item{Keep the \lstinline|for|\index{for} statement simple.

Avoid using the comma operator in the initialisation and update clause.}
\item{Keep the loop body short.

Use a label\index{label} for longer loops.}
\item{Use the \lstinline|for|\index{for} loop variant that fits best to your current problem.

The enhanced \lstinline|for|\index{for!enhanced} is not generally better than the basic variant, it just matches better with some iteration tasks.}
\end{enumerate}

\lstinline|for| loops are discussed in more detail in chapter \tqvref{sec:TheForStatement}.

%==============================================================================
% End of file!!
%==============================================================================
