\chapter{Summary}
Several rules and recommendation in this document, in particular regarding how to apply comments and how to write comments, but also those about to write a method, do require significant additional typing. Java in general has the reputation to be too verbose, and this coding conventions will even add to that.

But from my experience I found that verbosity is your friend!

And if you are afraid of the typing work: as I said already earlier (see chapter \tqref{sec:LengthOfNamesAndUseOfAbbreviations}), you should learn typewriting! For my understanding, someone who does not reach at least 100~CPM\footnote{CPM = “Characters per minute”, or, in German: „Anschläge pro Minute“} should look for a job outside of software development!

Writing code is not the only area where you would benefit from mastering that important skill; it will also help you to write all the other stuff you have to deliver in addition to your code (documentation, meeting notes, emails, specification documents, …).

\section{The “Zen of Java”}
In chapter \tqfullvref{sec:OtherProgrammingLanguages}, I quoted the “Zen of Python”\autocite{WIKIPEDIA:ZenOfPython,PYTHON_ORG_MAILING_LIST:ThePythonWay} and I referred to that set of rules several times in this document.

I did this because the “Zen of Python” is a \textit{nearly} generic coding guideline in a nutshell, valid for most, if not all programming languages.

Ok, some of the statements may need an explanation and/or some samples, and I hope that this document will give it to you.

But to get your Coding Guardrails in a nutshell, you can take the “Zen of Java” (stolen from the “Zen of Python” and slightly amended …) from here:
\begin{enumerate}[nosep]
\item Beautiful is better than ugly.
\item Explicit is better than implicit and verbosity is your friend.
\item Simple is better than complex.
\item Complex is better than complicated.
\item Flat is better than nested.
\item Sparse is better than dense.
\item Readability counts.
\item Special cases aren't special enough to break the rules …
\item … although practicality beats purity.
\item Errors should never pass silently and Exceptions may never be swallowed …
\item … unless explicitly silenced.
\item In the face of ambiguity, refuse the temptation to guess.
\item There should be one – and preferably only one – obvious way to do\\ it …
\item … although that way may not be obvious at first.
\item Now is better than never …
\item … although never is often better than \textit{right now}.
\item If the implementation is hard to explain, it's a bad idea.
\item If the implementation is easy to explain, it may be a good idea.
\item Think about your future self and add a comment.
\end{enumerate}

\section{How to use this Document?}
First, you can use this document as is! It should work for you, as its predecessors worked for a bunch of project teams.

But you can also shape it closer to your particular needs, if you want! Replace the examples and the references to my libraries by some that fits better to your project or your company.

This document was compiled using \TeX/\LaTeX\footnote{In case it is relevant for you, I used ‘TexMaker’ as the editor for the \LaTeX~sources. I run it on Mac and Linux, but the software is also available for Windows; it can be downloaded from this location: \href{https://www.xm1math.net/texmaker/download.html}{https://www.xm1math.net/texmaker/download.html}.}, and its source can be found on GitHub at \href{https://github.com/tquadrat/documents}{https://github.com/tquadrat/documents}. Clone the repository and startover!
